---
title: "Draft: Results"
---

Specific data on the energy usage for the Dauphin Island power plant could not be found, so the Hog Bayou Energy Center data was used to gather the results of the study. This plant sits alongside the Mobile River which feeds into the gulf, and is about 30 miles North of Dauphin Island. Data depicting the water temperature near the island was taken from NOAA, where a max time period of one year with one-hour intervals could be visualized on a graph or in table form. The most recent one-year time frame that had a full year's worth of data was from March 9th, 2022 to March 8th, 2023.

::: callout-note

You may want to combine data from several stations around Dauphin Island. They moved the stations, but there is a much longer period of data, and you should have that in order to get a sense of what has happened over the past 30 years wrt global warming.

:::

```{r}
library(tidyverse)
Dauphin_Island <- read.csv("dauphin_island.csv")
```

```{r}
eighty_five <-Dauphin_Island %>%
  filter(Water.Temp...F. > 85)

nrow(eighty_five)
```

The total number of hours in our time frame where water temperature was measured is 8759 hours. If we take the proportion of time where the water temperature in the gulf is above 85 degrees Fahrenheit, we would get a value of 0.2017 (20.17%).

```{r}
high_temp <- Dauphin_Island %>%
  filter(Water.Temp...F. > 85) %>%
  mutate(Month = month(Date, label = TRUE))
month_freq <- high_temp %>%
  count(Month) %>%
  arrange(desc(n))
print(month_freq)
```

It was reasonable for us to suspect that the water temperature would exceed 85 degrees the most often in the summer months, but it is still important that we narrow down exactly when during the summer that the plant will have to operate at reduced power. From the table, we can see that the plant will have to operate at reduced capacity in July, followed by June and August. According to @climate2025, the 10 warmest years in history have all been in the last decade, and the Earth is showing no signs of cooling down.


::: callout-note

Ok, I'd expect that you'd do something like regress water temperature on electrical demand to determine whether there's a linear relationship, or make a scatterplot by month of demand x temperature, or something of that sort.

You need a lot more analysis here to discuss, so that you have results to discuss that are useful and answer the question of interest. 

Also consider how to make your R output presentation ready -- can you make a table? A chart? You do not want raw output when your work is presented to a supervisor formally. 

:::
