---
title: "Results"
---

Because direct energy demand for Dauphin Island was unavailable, data from the Hog Bayou Energy Center (30 miles North) was used instead. This plant shares similar environmental conditions as it sits on the same river system feeding into the Gulf.  

```{r}
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
library(broom)
library(kableExtra)
Dauphin_Island <- read.csv("dauphin_island.csv")
energy_demand <- read.csv("energy_demand.csv")
```

## Water Temperature Analysis

```{r}
#| echo: false
#| message: false
#| warning: false

eighty_five <-Dauphin_Island %>%
  filter(Water.Temp...F. > 85)

```

The total number of hours in our time frame where water temperature exceeded 85 degrees is `r nrow(eighty_five)` hours out of 8759. If we take the proportion of time where the water temperature in the gulf is above 85 degrees Fahrenheit, we would get a value of 0.2017 (20.17%).

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: Figure-1

high_temp <- Dauphin_Island %>%
  filter(Water.Temp...F. > 85) %>%
  mutate(Month = month(Date, label = TRUE))
month_freq <- high_temp %>%
  count(Month) %>%
  arrange(desc(n))

month_freq |>
  rename(Month = Month, `Hours Above 85 degrees Fahrenheit` = n) |>
  kable(caption = "Frequency of High Water Temperature Occurrences by Month") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "left")
```

```{r}
#| echo: false
#| message: false
#| warning: false

water <- Dauphin_Island |>
  mutate(Date = as.Date(Date)) |>
  group_by(Date) |>
  summarize(avg_water_temp = mean(as.numeric(Water.Temp...F.), na.rm = TRUE))

energy <- energy_demand |>
  mutate(Date = as.Date(Timestamp..Hour.Ending., format = "%m/%d/%Y"))

combined <- inner_join(water, energy, by = "Date")
```

It was reasonable for us to suspect that the water temperature would exceed 85 degrees most often in the summer months, but it is still important that we narrow down exactly when during the summer that the plant will have to operate at reduced power. From the table, we can see that the plant will have to operate at reduced capacity in July, followed by June and August. No other months recorded temperatures above 85 degrees. According to @climate2025, the 10 warmest years in history have all been in the last decade, and the Earth is showing no signs of cooling down.

```{r}
#| echo: false
#| message: false
#| warning: false

montly_summary <- combined |>
  mutate(Month = month(Date, label = TRUE, abbr = TRUE)) |>
  group_by(Month) |>
  summarize(avg_temp = mean(avg_water_temp, na.rm = TRUE))

ggplot(montly_summary, aes(x=Month, y=avg_temp, fill = avg_temp)) +
  geom_col() +
  scale_fill_gradient(low = "blue", high = "red") +
  labs(title = "Average Water Temperature (F)", x= "Month", y = "Temp (F)") +
  theme_minimal()

```

The bar chart above showcases the average water temperatures observed in each month. It supports the table by showing us that from June to Spetember, we are likely to see consistent observations above 85 degrees.

## Regression and Residuals

```{r}
#| echo: false
#| message: false
#| warning: false

water <- Dauphin_Island |>
  mutate(Date = as.Date(Date)) |>
  group_by(Date) |>
  summarize(avg_water_temp = mean(as.numeric(Water.Temp...F.), na.rm = TRUE))

energy <- energy_demand |>
  mutate(Date = as.Date(Timestamp..Hour.Ending., format = "%m/%d/%Y"))

combined <- inner_join(water, energy, by = "Date")

ggplot(combined, aes(x = avg_water_temp, y = SOCO.Demand..MWh.)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "firebrick", se = TRUE) +
  labs(title = "Water Temperature vs. Energy Demand", x = "Average Water Temperature (F)", y = "Energy Demand (MWh)") +
  theme_minimal()
```

```{r}
#| echo: false
#| message: false
#| warning: false

model <- lm(SOCO.Demand..MWh. ~ avg_water_temp, data = combined)

tidy(model) |>
  kable(caption = "Linear Regression Results: Water Temperature vs. Energy Demand", digits = 30)
model_stats <- glance(model)
cat("The R-squared value for this model is:", round(model_stats$r.squared, 4))
```

From the table and the linear regression model, we can see that for every degree Fahrenheit that the water temperature increases, the energy demand is predicted to increase by 4,877 MWh, with an intercept starting at 289,619 MWh if the water temperature was 0 degrees. 

```{r}
#| echo: false
#| message: false
#| warning: false

combined <-combined |>
  mutate(residuals = resid(model), predicted = predict(model))

combined$residuals <- resid(model)

ggplot(combined, aes(x = avg_water_temp, y = residuals)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residual Plot: Water Temperature", x = "Average Water Temperature (F)", y = "Residuals (Error)") +
  theme_minimal()

ggplot(combined, aes(x = Date, y = residuals)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Residual Plot: Over Time", x = "Date", y = "Residuals (Error)") + 
  theme_minimal()

ggplot(combined, aes(x = SOCO.Demand..MWh., y = residuals)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Residual Plot: Energy Demand", x = "SOCO Demand (MWh)", y = "Residuals (Error)") + 
  theme_minimal()

```

The residual plot showcasing water temperature follows a distinct U-shape, proving that the linear model is biased and is incapable of following the trend of the data. It **overestimates** the demand in mild weather and **underestimates** it during extreme cold or heat. The residual plot with respect to time follows a wave shape, indicating seasonality. It shows that the model's errors aren't random, but they change predictably depending on the time of year. Finally, the residual plot with respect to energy demand follows an upward trend. It shows that as the energy demand grows, the model becomes less accurate.

```{r}
#| echo: false
#| message: false
#| warning: false

water <- Dauphin_Island |>
  mutate(Date = as.Date(Date)) |>
  group_by(Date) |>
  summarize(avg_water_temp = mean(as.numeric(Water.Temp...F.), na.rm = TRUE))

energy <- energy_demand |>
  mutate(Date = as.Date(Timestamp..Hour.Ending., format = "%m/%d/%Y"))

combined <- inner_join(water, energy, by = "Date")

ggplot(combined, aes(x = avg_water_temp, y = SOCO.Demand..MWh.)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", formula = y ~ poly(x,2), color = "firebrick", se = TRUE) +
  labs(title = "Water Temperature vs. Energy Demand", x = "Average Water Temperature (F)", y = "Energy Demand (MWh)") +
  theme_minimal()
```

```{r}
library(knitr)
library(kableExtra)

model_comparison <- data.frame(Model = c("Linear Regression", "Polynomial Regression (Quadratic)"),
                               `Adj. R-Squared` = c(0.283, 0.702),
                               `Model Utility` = c("Poor: Errors in extreme weather", "Strong: Captures extreme weather spikes"))

kable(model_comparison, align = "llc", booktabs = TRUE) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) |>
  row_spec(2, bold = TRUE)
```

The linear regression model gave an R-squared value of 0.2831, while the polynomial regression gave an R-squared of 0.7. This shows that a simple linear regression model only explained about **28.3%** of the variance in energy demand, while the polynomial regression explained about **70%** of the variance. This is important because the linear regression failed to show the extreme values in the colder months, while the polynomial regression captures the true trend of the data, both in extreme heat and cold.